**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-02-18 doc/paper/rebuttal_v1.md

Because the manuscript has been nearly entirely rewritten, we did not think it was appropriate to submit a version of the new manuscript that indicated the changes - the entire manuscript was changed. We appreciate the insights and suggestions from the original reviewers and apologize for taking so long to resubmit the manuscript. We took their advice to heart and regenerated the data using the P6C4 PacBio chemistry.

@@

**Lines 279-282: “We found that the frequency of the most abundant 1-nt error paralleled the number of sequences. There were two sequences in the V4 dataset that occurred 76 times, one sequence in the V1-V5 dataset that occurred 30 times, and one sequence in the V3-V5 dataset that occurred 17 times.” Are there any general characteristics of the most abundant errors that could provide insight into why certain errors tend to occur? e.g. high AT content or high nucleotide diversity overall (or in a particular sliding window)?**

Unfortunately, we were not able to identify any signatures beyond the various insertion, deletion, and substitution biases in the overall data set.

**Figure 2 is bit difficult to follow with this “numbering” scheme, suggest to change to colored lines.**

Part of the design of this figure was that the regions generally overlap on top of each other indicating that the effectiveness of the methods is reproducible. We have replaced the numbers with colored symbols and increased the horizontal jitter.

### 2016-02-18 Schloss_PacBio16S_PeerJ_2016.Rmd
Add introductory paragraph

***Data analysis.***
All sequencing data were curated using mothur (v1.36)[@Schloss2009a] and analyzed using the R programming language [@R2016]. The raw data can be obtained from the Sequence Read Archive at NCBI under accession SRP051686, which are associated with BioProject PRJNXXXXX. Several specific features were incorporated into mothur to facilitate the analysis of PacBio sequence data. First, because non-ambiguous base calls are assigned to Phred quality scores of zero, the consensus fastq files were parsed so that scores of zero were interpreted as corresponding to an ambiguous base call (i.e. N) in the fastq.info command using the pacbio=T option. Second, because the consensus sequence can be generated in the forward and reverse complement orientations, a checkorient option was added to the trim.seqs command in order to identify the proper orientation. These features were incorporated into mothur v.1.30. Because chimeric molecules can be generated during PCR and would artificially inflate the sequencing error, it was necessary to remove these data prior to assessing the error rate. Because we knew the true sequences for the strains in the mock community we could calculate all possible chimeras between strains in the mock community (*in silico* chimeras). If a sequence read was 3 or more nucleotides more similar to an *in silico* chimera than it was to a non-chimeric reference sequence, it was classified as a chimera and removed from further consideration. Identification of *in silico* chimeras and calculation of sequencing error rates was performed using the seq.error command in mothur [@Schloss2011a]. *De novo* chimera detection was also performed on the mock and other sequence data using the abundance-based algorithm implemented in UCHIME [@Edgar2011]. Sequences sequences were aligned against a SILVA-based reference alignment [@Pruesse2007] using a profile-based aligner [@Schloss2009b] and were classified against the SILVA (v123) [@Pruesse2007], RDP (v10)[@Cole2013], and greengenes (v13_8_99)[@Werner2011] reference taxonomies  using a negative Bayesian classifier implemented within mothur [@Wang2007]. Sequences were assigned to operational taxonomic units using the average neighbor clustering algorithm with a 3% distance threshold [@Schloss2011b]. Detailed methods including this paper as an R markdown file are available as a public online repository (http://github.com/SchlossLab/Schloss_PacBio16S_PeerJ_2016).

>>>>

***Data analysis.***
All sequencing data were curated using mothur (v1.36)[@Schloss2009a] and analyzed using the R programming language [@R2016]. The raw data can be obtained from the Sequence Read Archive at NCBI under accession SRP051686, which are associated with BioProject PRJNA271568. This accession and bioproject also contain data from the same samples sequenced using the P4-C2 chemistry. Several specific features were incorporated into mothur to facilitate the analysis of PacBio sequence data. First, because non-ambiguous base calls are assigned to Phred quality scores of zero, the consensus fastq files were parsed so that scores of zero were interpreted as corresponding to an ambiguous base call (i.e. N) in the fastq.info command using the pacbio=T option. Second, because the consensus sequence can be generated in the forward and reverse complement orientations, a checkorient option was added to the trim.seqs command in order to identify the proper orientation. These features were incorporated into mothur v.1.30. Because chimeric molecules can be generated during PCR and would artificially inflate the sequencing error, it was necessary to remove these data prior to assessing the error rate. Because we knew the true sequences for the strains in the mock community we could calculate all possible chimeras between strains in the mock community (*in silico* chimeras). If a sequence read was 3 or more nucleotides more similar to an *in silico* chimera than it was to a non-chimeric reference sequence, it was classified as a chimera and removed from further consideration. Identification of *in silico* chimeras and calculation of sequencing error rates was performed using the seq.error command in mothur [@Schloss2011a]. *De novo* chimera detection was also performed on the mock and other sequence data using the abundance-based algorithm implemented in UCHIME [@Edgar2011]. Sequences sequences were aligned against a SILVA-based reference alignment [@Pruesse2007] using a profile-based aligner [@Schloss2009b] and were classified against the SILVA (v123) [@Pruesse2007], RDP (v10)[@Cole2013], and greengenes (v13_8_99)[@Werner2011] reference taxonomies  using a negative Bayesian classifier implemented within mothur [@Wang2007]. Sequences were assigned to operational taxonomic units using the average neighbor clustering algorithm with a 3% distance threshold [@Schloss2011b]. Detailed methods including this paper as an R markdown file are available as a public online repository (http://github.com/SchlossLab/Schloss_PacBio16S_PeerJ_2016).
